---
output: 
  github_document: 
    toc: true
params: 
  weekday: Monday
editor_options: 
  chunk_output_type: inline
---

---
title: `r params$weekday`  
author: "Shih-Ni Prim"  
date: "2020-10-07"  
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
```

```{r, echo = FALSE, eval = FALSE}
# dayz <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
# paramz <- lapply(dayz, FUN = function(x){list(weekday = x)})
# output_file <- paste0("Report-", dayz, ".md")
# reports <- tibble(output_file, paramz)
# 
# library(rmarkdown)
# 
# apply(reports, MARGIN = 1,
#       FUN = function(x){
#           render(input = "analysis.Rmd",
#                  output_format = "github_document",
#                  output_file = x[[1]],
#                  params = x[[2]])
#           })

rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Sunday.md", params = list(weekday = "Sunday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Monday.md", params = list(weekday = "Monday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Tuesday.md", params = list(weekday = "Tuesday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Wednesday.md", params = list(weekday = "Wednesday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Thursday.md", params = list(weekday = "Thursday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Friday.md", params = list(weekday = "Friday"))
rmarkdown::render(input = "analysis.Rmd", output_format = "github_document", output_file = "Report-Saturday.md", params = list(weekday = "Saturday"))
```

This is `r params$weekday`'s analysis.  

## Introduction  

This about bike share shows the information about each day--hours, temperature, humidity, weekday, holiday/workday or not, etc.  

Since the current analysis is on `r params$weekday`, we first find the corresponding value for it.  

```{r}
set.seed(7777)
i <- 0:6
dayz <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
df <- as.data.frame(cbind(i, dayz))
weekdayNum <- df$i[df$dayz == params$weekday]
print(weekdayNum)
```


## Data  

Now we read in the data.  

```{r}
bikes <- read_csv("../Bike-Sharing-Dataset/hour.csv")
# head(bikes)
analysis <- bikes %>% filter(weekday == weekdayNum) %>% select(-casual, -registered) %>% select(dteday, weekday, everything()) 
# head(analysis)
```

## Splitting up to training and test sets  

```{r}
train <- sample(1:nrow(analysis), size = nrow(analysis)*0.7)
test <- setdiff(1:nrow(analysis), train)

bikeTrain <- analysis[train,]
bikeTest <- analysis[test,]
```

```{r}
ggplot(bikeTrain, mapping = aes(x = cnt)) + geom_histogram()
ggplot(bikeTrain, aes(x = hr, y = cnt)) + geom_point() + geom_jitter()
ggplot(bikeTrain, aes(x = yr, y = cnt)) + geom_boxplot(aes(group = yr))
cor(bikeTrain$mnth, bikeTrain$season)
ggplot(bikeTrain, aes(x = weathersit, y = windspeed)) + geom_jitter()
cor(bikeTrain$holiday, bikeTrain$workingday)
cor(bikeTrain$temp, bikeTrain$atemp)
var(bikeTrain$holiday)
var(bikeTrain$workingday)
```

`mnth` and `season` as well as `workingday` and `holiday` are highly correlated. Also, `instant` and `dteday` are for record-keeping. The variance of `workingday` and `holiday` is 0 or too small. Thus, we decide to keep these as the predictors: `season`, `yr`, `hr`, `weathersit`, `atemp`, `hum`, and `windspeed`.  

```{r}
bikeTrain <- select(bikeTrain, season, yr, hr, weathersit, atemp, hum, windspeed, cnt)
bikeTest <- select(bikeTest, season, yr, hr, weathersit, atemp, hum, windspeed, cnt)
```


## Fitting models  


### Regression tree  

```{r}
modelLookup("rpart")

bikeTree <- train(cnt ~ ., data = bikeTrain, method = "rpart", trControl = trainControl(method = "cv", number = 10), tuneLength = 10)

predTree <- predict(bikeTree, newdata = bikeTest)
treeRMSE <- sqrt(mean((predTree - bikeTest$cnt)^2))
postResample(predTree, bikeTest$cnt)
```

### Boosted Tree  

```{r}
modelLookup("gbm")

boostedBike <- train(cnt ~  season + yr + hr + weathersit + atemp + hum + windspeed, data = bikeTrain, method = "gbm", preProcess = c("center", "scale"), trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3), tuneLength = 5, verbose = FALSE)
predBoostedBike <- predict(boostedBike, newdata = select(bikeTest, -cnt))
boostedRMSE <- sqrt(mean((predBoostedBike - bikeTest$cnt)^2))
postResample(predBoostedBike, bikeTest$cnt)
```

### Comparison  

```{r}
comparison <- data.frame(treeRMSE, boostedRMSE)
colnames(comparison) <- c("Regression Tree", "Boosted Tree")
rownames(comparison) <- c("RMSE")
print(comparison)
```



```{r}
model <- function(x, y){
  if (x > y) {
    final <- c("boosted tree")
  }
  else {
    final <- c("regression tree")
  }
  return(final)
}

# model(treeRMSE, boostedRMSE)
```

From the output, we can conclude that `r model(treeRMSE, boostedRMSE)` is the better model.  
